> We show a R function and a Rcpp function in this file.

## Function 1
We realize an algorithm in the paper 'Sparse Reduced-Rank Regression for Simultaneous Dimension Reduction and Variable Selection'("https://www.tandfonline.com/doi/full/10.1080/01621459.2012.734178"). The algorithom is called 'Variational Method', which solves the optimization problem:
$$min_{A,B} \Vert \mathbf{Y}-\mathbf{XBA^{T}} \Vert^{2} + \sum_{i=1}^{p} \lambda_{i}\Vert \mathbf{B}^{i} \Vert$$ such that 
$\mathbf{A^{T}A = I}$. It iteratively optimize with respect to $\mathbf{A}$ and $\mathbf{B}$.

```{r}
# Data Initialization

knitr::opts_chunk$set(echo = TRUE)
library(MASS)

set.seed(22004)
n = 20
p = 6
q = 6
p0 = 1
r = 2
rho.x = 0.1
rho.e = 0.1
lambda <- rep(0.5, p)

sample.generator <- function(n, p, p0, q, r, rho.x, rho.e){
  X <- matrix(0, nrow = n, ncol = p)
  Y <- E <- matrix(0, nrow = n, ncol = q)
  B <- matrix(c(rnorm(p0*r),rep(0,(p-p0)*r)), nrow = p, ncol = r, byrow = TRUE)
  A <- matrix(rnorm(q*r), nrow = q, ncol = r)
  C <- B%*%t(A)
  
  sigma.x <- matrix(rho.x, nrow = p, ncol = p)
  sigma.x <- sigma.x+diag(1-rho.x, p)
  X <- mvrnorm(n, mu = rep(0,p), Sigma = sigma.x)
  
  sigma.e <- matrix(rho.e, nrow = q, ncol = q)
  sigma.e <- sigma.e+diag(1-rho.e, q)
  sgm2 <- sum(diag(t(C)%*%sigma.x%*%C))/q
  E <- mvrnorm(n, mu = rep(0,q), Sigma = sgm2*sigma.e)
  
  Y <- X%*%C+E
  
  return(list(X=X,Y=Y, E=E, C=C, B=B, A=A))
}

X <- sample.generator(n, p, p0, q, r, rho.x, rho.e)$X
Y <- sample.generator(n, p, p0, q, r, rho.x, rho.e)$Y
E <- sample.generator(n, p, p0, q, r, rho.x, rho.e)$E
C <- sample.generator(n, p, p0, q, r, rho.x, rho.e)$C
A <- sample.generator(n, p, p0, q, r, rho.x, rho.e)$A
```

---
title: "Introduction to StatComp22004"
author: "Dingyun Zhang"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to StatComp22004}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---


```{r}
# Variational Method
SRRR.variational <- function(X, Y, lambda, r){
  obj <- function(Y, X, B, A, lambda){
    norm(Y-X%*%B%*%t(A),type = "F")^2 + sum(lambda*apply(B,1,function(x) sqrt(sum(x^2))))
    }
  
  n <- nrow(X)
  p <- ncol(X)
  q <- ncol(Y)
  Sxx <- t(X)%*%X/n
  Sxy <- t(X)%*%Y/n
  Syx <- t(Y)%*%X/n
  V <- eigen(Syx%*%solve(Sxx)%*%Sxy,symmetric = TRUE)$vector[,1:r]
  A <- V
  B <- solve(Sxx)%*%Sxy%*%V
  A.hat <- matrix(numeric(q*r), nrow = q, ncol = r)
  B.hat <- matrix(numeric(p*r), nrow = p, ncol = r)
  
  mu <- numeric(p)
  
  while(abs(obj(Y,X,B.hat,A.hat,lambda)-obj(Y,X,B,A,lambda))>1e-5){
    A.hat <- A
    svd.yxb <- svd(t(Y)%*%X%*%B)
    u <- svd.yxb$u
    v <- svd.yxb$v
    A <- u%*%t(v)
    
    while(norm(B-B.hat,type = "F")>1e-3){
      B.hat <- B
      mu <- apply(B,1,function(x) 1/sqrt(sum(x^2)))
      B <- ginv(t(X)%*%X+diag(lambda*mu)/2)%*%t(X)%*%Y%*%A
    }
  }
  
  return(list(A=A,B=B))
}
```

```{r}
# Test
SRRR.variational(X, Y, lambda, r)
```

## Function 2

We realize stochastic gradient descent(SGD) algorithm in Rcpp.

```{c,eval=FALSE}
#include <Rcpp.h>
using namespace Rcpp;
// [[Rcpp::export]]
NumericVector SGD_C(IntegerMatrix matrix, NumericVector result, NumericVector theta, float loss){
  for(int i =0 ;i<100 && loss>0.001; ++i){
    float error_sum = 0.0;
    int j=i%4;//j=0,1,2,3
    float h = 0.0;
    for(int k=0; k<2; ++k){
      h += matrix(j,k)*theta[k];
      }
    error_sum = result[j]-h;
    for(int k=0;k<2;++k){
      theta[k] += 0.01*(error_sum)*matrix(j,k);
      }
    printf("parameters:%f,%f\n",theta[0],theta[1]);
    float loss = 0.0;
    for(int j = 0;j<4;++j){
      float sum=0.0;
      for(int k = 0;k<2;++k){
        sum += matrix(j,k)*theta[k];
        }
      loss += (sum-result[j])*(sum-result[j]);
      }
    printf("iter:%d\n",i);
    printf("loss:%f\n",loss);
    }
  return theta;
}
```

```{r,warning=FALSE}
# Test
library(Rcpp)

# Data Initialization
matrix <- matrix(c(1,4,2,5,5,1,4,2), nrow = 4, ncol = 2)
result <- c(19,26,19,20)
theta <- c(2,5)
loss <- 1000.0

sourceCpp('C:/Users/10313/Desktop/StatisticalComputing/HW/HW11/SGD_C.cpp')
SGD_C(matrix, result, theta, loss)

```